---
title: "Super document"
author: "Winnie the Pooh"
output:
  html_document: default
  word_document: default
lang: en-GB
editor_options: 
  chunk_output_type: console
---

Let's start our small research!

# Load the data

First, we attach some packages:

```{r, message=FALSE, echo=FALSE}
library(tidyverse) # working with data
library(rio) # import / export
library(texreg) # table for many regressions
library(lmtest) # tests in linear models
library(car) # companion for applied regression
```

Then we load the data:
```{r, echo=FALSE}
flats <- import("~/Downloads/flats_moscow.txt")
```

First plot

```{r, echo=FALSE}
qplot(data = flats, x = totsp, y = price) +
  xlab("Total span, square meters") +
  ylab("Price, thouthands of dollars") +
  ggtitle("Flat prices in Moscow, 2007")
```

Our dataset contains `r nrow(flats)` observations.

According to latest scientific discoveries logarithm of two is equal to `r log(2)`.

Let's estimate some models:
```{r}
model_a <- lm(data = flats, price ~ totsp)
model_b <- lm(data = flats, 
    price ~ totsp + livesp + kitsp)
model_c <- lm(data = flats, 
    price ~ totsp + livesp + kitsp + brick)
```


just the result
```{r, eval=FALSE}
summary(model_a)
summary(model_b)
summary(model_c)
```

Now the same results in a pretty table

```{r, results="asis"}
htmlreg(list(model_a, model_b, model_c))
```

# Model comparison

I am happy and I would like to compare model A and model C:

```{r}
waldtest(model_a, model_c)
```

As p-value is less than $\alpha=0.05$, short model (model a) is rejected.

## Is the pricing of brick and non-brick flats similar?

```{r}
reg_whole <- lm(data = flats, 
     price ~ totsp + livesp + kitsp)
```

summary for regression
```{r}
summary(reg_whole)
```


Extract Residual Sum of Squares
```{r}
deviance(reg_whole)
```

Look inside complex object with all results:
```{r}
summ_whole <- summary(reg_whole)
r2_whole <- summ_whole[["r.squared"]]
```

For the model on the whole dataset the $R^2$ is equal to `r r2_whole` and RSS is equal to `r deviance(reg_whole)`.

Now we filter data frame:
```{r}
flats_brick <- filter(flats, brick == 1)
flats_non_brick <- filter(flats, brick == 0)
```

Two regressions on subsets:
```{r}
reg_brick <- lm(data = flats_brick, 
     price ~ totsp + livesp + kitsp)
reg_non_brick <- lm(data = flats_non_brick, 
     price ~ totsp + livesp + kitsp)
```

Put all results into one table
```{r, results="asis"}
htmlreg(list(reg_whole, reg_brick, reg_non_brick))
```

Manually to F-test
```{r}
RSS_R <- deviance(reg_whole)
RSS_UR <- deviance(reg_brick) + deviance(reg_non_brick)
n <- nrow(flats)
F_obs <- (RSS_R - RSS_UR) / 4 / (RSS_UR / (n - 8))
F_crit <- qf(0.95, df1 = 4, df2 = n - 8)
```

In our test the observed value of F-statistic is `r F_obs`, while the critical value is `r F_crit`. So the null hyphotesis of the same model for each subset is rejected. 


Another way! We may estimate one unrestricted model on the whole dataset instead of two models on subsets.

```{r}
reg_unrestricted <- lm(data = flats, 
                price ~  brick * (1 + totsp + livesp + kitsp))
summary(reg_unrestricted)
```

Second way of doing F-test:
```{r}
waldtest(reg_whole, reg_unrestricted)
```

The p-value is too small, so the null-hypothesis of same model for both subsets is rejected. Models for brick and non-brick flats are different.


# Dummy variables for encoding of cathegorical variables

Wrong method!!!
```{r}
model_wrong <- lm(data = flats, 
                  price ~ totsp + livesp + kitsp + code)
summary(model_wrong)
```

Good method!!!
```{r}
flats2 <- mutate(flats, code = factor(code))
glimpse(flats2)
model_good <- lm(data = flats2, 
                  price ~ totsp + livesp + kitsp + code)
summary(model_good)
```

Test hypothesis that zone 3 is similar to zone 8.
```{r}
flats2 <- mutate(flats2, 
                 code_alt = ifelse(code == "8", "3", code))
glimpse(flats2)
model_rest <- lm(data = flats2, 
                  price ~ totsp + livesp + kitsp + code_alt)
RSS_R <- deviance(model_rest)
RSS_UR <- deviance(model_good)
F_obs <- (RSS_R - RSS_UR) / 1 / (RSS_UR / (n - 11))
F_obs
F_crit <- qf(0.95, df1 = 1, df2 = n - 11)
F_crit
```


Test hypothesis that zone 3 is similar to zone 8.
```{r}
linearHypothesis(model_good, "code3 = code8")
```

Test hypothesis that zone 3 is similar to zone 8 and zone 1 simultaneously.
```{r}
linearHypothesis(model_good, c("code3 = code8", "code3 = 0"))
```










