---
title: "Безумно важное исследование"
author: "Winnie the Pooh"
date: "19 декабря 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Обзор литературы / Literature review

Ещё питекантропы говорили о важности эконометрики
Long time ago...
```{r, message=FALSE, echo=TRUE}
library(tidyverse) # data manipulations
library(lmtest) # model tests
library(texreg) # tables for model comparison
```

Вставка блока кода через Code — Insert chunk
Insert code chunk using Code — Insert chunk

График набора данных про бриллианты в студию!
Scatterplot of dataset!
```{r}
qplot(data = diamonds, x = carat, y = price)
```

# Основная часть (main part)

Начало нашего набора данных
Top rows of `diamonds` dataset
```{r}
head(diamonds)
```

А не оценить ли благородному дону пару регрессий?
Estimate two models. Dependent variable is to the left of tilde and predictors — to the right of tilde.
```{r}
model_a = lm(data = diamonds, price ~ carat)
model_b = lm(data = diamonds, price ~ carat + x + y + z)
summary(model_b)
```

Сравним рядом две модели в одной табличке
Compare two models in one table
```{r, results='asis'}
htmlreg(list(model_a, model_b))
# screenreg(list(model_a, model_b))
```

Сравним с помощью F-теста две вложенных модели
Compare two nested models using F-test:
```{r}
waldtest(model_a, model_b)
```

Построим доверительный интервал для коэффициентов:
We calculate confidence interval for true coefficients:
```{r}
confint(model_b, level = 0.95)
```

```{r}
library(Ecdat) # a lot of data sets for studies :)
library(skimr) # descriptive statistics
```

* Look at the documentation for `Icecream` dataset.
Just `?...`
* Calculate descriptive stats for `Icecream` dataset. 
Hint: use `skim(...)` funciton :) 
* Plot `temp` (temperature) against `cons` (icecream consumtion).
* Estimate regression of `cons` (consumption of icecream)
on `income`, `price` and `temp` (temperature). This is Model A. 
* Calculate 90% confidence interval for each coefficient
* Compare Model A with simplier Model B: 
`cons` consumption depends only on `temp` temperature.
Do F-test and a table for two models. Which one is better?

# Load your own dataset :)

```{r}
library(rio)
mydata = import("~/Downloads/demo.xlsx")
mydata
skim(mydata) # descriptive statistics
```

If you don't know where the file is :)
In console run the command `file.choose()`



