---
title: "Привет"
author: "Winnie the Pooh"
date: "1/23/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

Let's attach some packages!
```{r}
library(rio) # imports data from various formats
library(tidyverse) # data manipulations and plots
library(lmtest) # test for linear models
library(skimr) # summary statistics
```

Load the dataset
```{r}
flats = import("~/Downloads/flats_moscow.txt")
```

If you don't remember even approximate location 
you may just run command `file.choose()` in console.

Look at the dataset
```{r}
glimpse(flats) # general info about dataset (n obs, n variables...)
skim(flats) # descriptive stats
head(flats) # beginning
tail(flats) # end of the dataset
```


Some quick plots!
```{r}
qplot(data = flats, x = log(price)) + 
  labs(xlab = "Log price 1000$",
  ylab = "Number of flats",
  title = "Price distribution")
```

Joint plot
```{r}
qplot(data = flats, 
      x = log(totsp), y = log(price)) + 
  labs(xlab = "Total span, square meters",
  ylab = "Price of flat in 1000$",
  title = "Moscow flats prices in 2009")
```

First model. Price depends on room area.
```{r}
model_a = lm(data = flats, price ~ totsp + livesp + kitsp)
summary(model_a)
```

Important: 1.29 1000$ is the price of balcony, bathroom etc

Let's introduce new variable `othersp`
```{r}
flats = mutate(flats, othersp = totsp - kitsp - livesp)
model_b = lm(data = flats, price ~ livesp + kitsp + othersp)
summary(model_b)
```

Calculate confidence intervals
```{r}
confint(model_b, level = 0.95)
```

Let's estimate more complex model
```{r}
model_c = lm(data = flats, 
    price ~ totsp + livesp + kitsp + brick + floor + walk)
summary(model_c)
```

Let's look at $RSS = \sum (y_i - \hat y_i)^2$:
```{r}
RSS_a = deviance(model_a) # RSS
RSS_c = deviance(model_c) # RSS
RSS_a
RSS_c
```

Let's add some junk variables:
```{r}
flats3 = mutate(flats, noise1 = rnorm(2040), 
                noise2 = rnorm(2040))
head(flats3)
model_d = lm(data = flats3, 
    price ~ totsp + livesp + kitsp + 
      brick + floor + walk + noise1 + noise2)
summary(model_d)
deviance(model_d)
```

Let's compare two models with F-test
```{r}
waldtest(model_c, model_d)
```

Let's find F critical :)
```{r}
F_crit = qf(0.95, df1 = 2, df2 = 2040 - 9)
F_crit
```

F observed (0.612) is less than F critical (3.00), so H_0 is not rejected.

Or:

P-value (0.5424) is bigger than significance level (alpha = 0.05), so H_0 is not rejected.

Let's compare model A and model C.
```{r}
waldtest(model_a, model_c)
```



