\documentclass[12pt]{article}

\usepackage{tikz} % картинки в tikz
\usepackage{microtype} % свешивание пунктуации

\usepackage{array} % для столбцов фиксированной ширины

\usepackage{indentfirst} % отступ в первом параграфе

\usepackage{sectsty} % для центрирования названий частей
\allsectionsfont{\centering}

\usepackage{amsmath, amssymb} % куча стандартных математических плюшек


\usepackage{comment}

\usepackage[top=2cm, left=1.2cm, right=1.2cm, bottom=2cm]{geometry} % размер текста на странице

\usepackage{lastpage} % чтобы узнать номер последней страницы

\usepackage{enumitem} % дополнительные плюшки для списков
%  например \begin{enumerate}[resume] позволяет продолжить нумерацию в новом списке
\usepackage{caption}


\usepackage{fancyhdr} % весёлые колонтитулы
\pagestyle{fancy}
\lhead{Econometrics}
\chead{}
\rhead{2020-06-25, final exam}
\lfoot{}
\cfoot{}
\rfoot{\thepage/\pageref{LastPage}}
\renewcommand{\headrulewidth}{0.4pt}
\renewcommand{\footrulewidth}{0.4pt}


\let\P\relax
\DeclareMathOperator{\P}{\mathbb{P}}

\usepackage{todonotes} % для вставки в документ заметок о том, что осталось сделать
% \todo{Здесь надо коэффициенты исправить}
% \missingfigure{Здесь будет Последний день Помпеи}
% \listoftodos --- печатает все поставленные \todo'шки


% более красивые таблицы
\usepackage{booktabs}
% заповеди из докупентации:
% 1. Не используйте вертикальные линни
% 2. Не используйте двойные линии
% 3. Единицы измерения - в шапку таблицы
% 4. Не сокращайте .1 вместо 0.1
% 5. Повторяющееся значение повторяйте, а не говорите "то же"



\usepackage{fontspec}
\usepackage{polyglossia}

\setmainlanguage{russian}
\setotherlanguages{english}

% download "Linux Libertine" fonts:
% http://www.linuxlibertine.org/index.php?id=91&L=1
\setmainfont{Linux Libertine O} % or Helvetica, Arial, Cambria
% why do we need \newfontfamily:
% http://tex.stackexchange.com/questions/91507/
\newfontfamily{\cyrillicfonttt}{Linux Libertine O}

\AddEnumerateCounter{\asbuk}{\russian@alph}{щ} % для списков с русскими буквами
%\setlist[enumerate, 2]{label=\asbuk*),ref=\asbuk*}

%% эконометрические сокращения
\DeclareMathOperator{\Cov}{Cov}
\DeclareMathOperator{\Corr}{Corr}
\DeclareMathOperator{\Var}{Var}
\DeclareMathOperator{\E}{E}
\def \hb{\hat{\beta}}
\def \hs{\hat{\sigma}}
\def \htheta{\hat{\theta}}
\def \s{\sigma}
\def \hy{\hat{y}}
\def \hY{\hat{Y}}
\def \v1{\vec{1}}
\def \e{\varepsilon}
\def \he{\hat{\e}}
\def \z{z}
\def \hVar{\widehat{\Var}}
\def \hCorr{\widehat{\Corr}}
\def \hCov{\widehat{\Cov}}
\def \cN{\mathcal{N}}


\begin{document}

\section*{Promo-code activation :)}

You have two options! 

You can try to solve all the proposed problems and submit your solutions as a pdf file.
In this case I will honestly grade them.

You can write a promo-code \#beta\_hot. 
In this case I will ignore your solutions and grade your exam as 4/10. 

\section*{Problems}

\begin{enumerate}

	\item James Bond estimates the model $y_t = \beta x_t + u_i$ with OLS.
	We know that $u_t$ are independent and identically distributed with unknown heteroskedasticity structure.
	We also know that $x_t = 1$ for odd $t$ and $x_t = 2$ for even $t$.
	The total number of observations $n$ is 600. 
  
	\begin{tabular}{ccc}
	  \toprule
		 & all odd $t$ & all even $t$ \\
	  \midrule
	  $\sum y_t$ & 300 & 200  \\
	  $\sum \hat u_t^2$ & 1000 & 800 \\
	  \bottomrule
  \end{tabular}
  
  
   \begin{enumerate}
	  \item Find $\hat\beta_{ols}$
	   \item Calculate ordinary and robust standard errors, $se(\hat\beta_{ols})$ and $se_{HC}(\hat\beta_{ols})$;
	   \item Construct 95\% ordinary and robust confidence intervals for $\beta$. 
   \end{enumerate}
  


	\item James Bond estimates the model $y_t = \beta x_t + u_i$ with OLS.
	We know that $u_t$ are normal $\cN(0; x_t \cdot \sigma^2)$ and independent.
	We also know that $x_t = 1$ for odd $t$ and $x_t = 2$ for even $t$.
	The total number of observations $n$ is even. 
	    \begin{enumerate}
      \item Find expected value and variance of $\hat \beta_{ols}$.
      \item Construct the most efficient unbiased linear estimator $\hat\beta_{best}$ and find its variance.
 	  \item Calculate probability limit of $\hat \beta_{ols}$. Is $\hat \beta_{ols}$ consistent?
    \end{enumerate}

 \item James Bond assumes that observations $y_i$ are independent and probability distribution for $y_i$ is given by the table

	 \begin{tabular}{cccc}
		 \toprule
		 $y_i$ & $y_i = 1$ & $y_i = 2$ & $y_i = 3$ \\
		 \midrule
		 probability & $a$ & $b$ & $1 - a-b$ \\
		 \bottomrule
	 \end{tabular}

	 James Bond has collected 400 observations: $120$ observations are equal to $1$, $80$ observations equal to $2$, and $200$ observations equal to $3$.

\begin{enumerate}
 \item Estimate unknown $a$ and $b$ using maximum likelihood;
 \item Construct 95\% confidence interval for $a$ and 95\% confidence interval for $a+b$.
 \item Using likelihood ratio statistic test the hypothesis $a=b$. You may assume that the critical value
 of the statistic is $4$ but you should clearly state the degrees of freedom. 
\end{enumerate}

\newpage
\item Using 400 observation James Bond estimated the logistic regression for probability
of passing econometrics exam using $x_i$ — number of study hours as predictor:
\[
\hat \P(\text{pass}) = \Lambda(0.15 - 0.025 x_i^2 +  0.9x_i),
\]

The estimate of coefficient covariance matrix is:
\[
\begin{pmatrix}
0.0004 & -0.0001 & 0 \\
-0.0001 & 0.0001 & 0 \\
0 & 0 & 0.0009 \\
\end{pmatrix}
\]

\begin{enumerate}
  \item Estimate the number of study hours that maximizes the probability of passing the exams.
  What is the maximal probability of passing the exam?
  \item Construct approximate 95\% confidence interval for the number of study hours that maximizes the probability of passing the exams.
  Hint: use Taylor expansion to obtain a linear function of $\hat\beta$.
\end{enumerate}

\item James Bond would like to check whether the econometrics final exam score, $y_i$, varies for
students that dinner at Khatchapuri. Once again $x_i$ is the number of study hours.

He estimates the same model using four datasets:

\begin{tabular}{cccc}
	\toprule
	dataset & equation & $RSS$ & observations \\
	\midrule
	all students & $\hat y_i = 40 + 2x_i$ & 14000 & 600 \\
	two or more Khatchapuri dinners & $\hat y_i = 60 + 4x_i$ & 4000 & 200 \\
	exactly one Khatchapuri dinner & $\hat y_i = 50 + 3x_i$ & 3000 & 200 \\
	no Khatchapuri dinner & $\hat y_i = 37 + 1.8x_i$ & 2000 & 200 \\
	\bottomrule
\end{tabular}

\begin{enumerate}
	\item Using appropriate test check whether the econometrics final exam score varies for three type of students.
	\item Which unique equation should James Bond estimate to obtain all given beta estimates at once?
\end{enumerate}


You may assume that $F_{critical}$ for $\alpha=0.05$ is approximately equal to $2.4$.
However you should clearly state the degrees of freedom.

\item The last problem :)

James Bond would like to estimate the model $y_i =  \beta x_i + u_i$,
where $y_i$ — is the econometrics final exam centered result, $x_i$ is the centered number of study hours.


The problem is that James has only data on self-reported $x^*_i = x_i \cdot f_i$ where
$f_i$ is the famous Russian exageration factor. We know that $x_i \sim \cN(0;9)$,
$u_i \sim \cN(0;\sigma^2)$, $f_i$ is uniform on $[0;4]$. 

James Bond also has estimates of the famous Russian exageration factor, $f^*_i$ for each student.
Here $f^*_i = f_i + \nu_i$ where $\nu_i \sim \cN(\delta, \sigma^2_{\nu})$. 
The parameters $\delta$ and $\sigma^2_{\nu}$ are unknown.

Variables $(x_i)$, $(u_i)$, $(\nu_i)$ and $(f_i)$ are independent. The dataset is a random sample. 



\begin{enumerate}
	\item Find the probability limit of $\hat\beta$ in the regression $\hat y_i = \hat\beta \cdot x^*_i$.
	\item Provide explicit formulas or algorithms for consistent estimators of all parameters: $\nu$, $\sigma^2_{\nu}$, $\beta$, $\sigma^2$.
\end{enumerate}





\end{enumerate}

\end{document}
