\documentclass[12pt]{article}

\usepackage{tikz} % картинки в tikz
\usepackage{microtype} % свешивание пунктуации

\usepackage{array} % для столбцов фиксированной ширины

\usepackage{indentfirst} % отступ в первом параграфе

\usepackage{sectsty} % для центрирования названий частей
\allsectionsfont{\centering}

\usepackage{amsmath, amssymb} % куча стандартных математических плюшек


\usepackage{comment}

\usepackage[top=2cm, left=1.2cm, right=1.2cm, bottom=2cm]{geometry} % размер текста на странице

\usepackage{lastpage} % чтобы узнать номер последней страницы

\usepackage{enumitem} % дополнительные плюшки для списков
%  например \begin{enumerate}[resume] позволяет продолжить нумерацию в новом списке
\usepackage{caption}


\usepackage{fancyhdr} % весёлые колонтитулы
\pagestyle{fancy}
\lhead{Econometrics}
\chead{}
\rhead{2019-09-23, final exam-retake}
\lfoot{}
\cfoot{}
\rfoot{\thepage/\pageref{LastPage}}
\renewcommand{\headrulewidth}{0.4pt}
\renewcommand{\footrulewidth}{0.4pt}


\let\P\relax
\DeclareMathOperator{\P}{\mathbb{P}}

\usepackage{todonotes} % для вставки в документ заметок о том, что осталось сделать
% \todo{Здесь надо коэффициенты исправить}
% \missingfigure{Здесь будет Последний день Помпеи}
% \listoftodos — печатает все поставленные \todo'шки


% более красивые таблицы
\usepackage{booktabs}
% заповеди из докупентации:
% 1. Не используйте вертикальные линни
% 2. Не используйте двойные линии
% 3. Единицы измерения — в шапку таблицы
% 4. Не сокращайте .1 вместо 0.1
% 5. Повторяющееся значение повторяйте, а не говорите "то же"



\usepackage{fontspec}
\usepackage{polyglossia}

\setmainlanguage{russian}
\setotherlanguages{english}

% download "Linux Libertine" fonts:
% http://www.linuxlibertine.org/index.php?id=91&L=1
\setmainfont{Linux Libertine O} % or Helvetica, Arial, Cambria
% why do we need \newfontfamily:
% http://tex.stackexchange.com/questions/91507/
\newfontfamily{\cyrillicfonttt}{Linux Libertine O}

\AddEnumerateCounter{\asbuk}{\russian@alph}{щ} % для списков с русскими буквами
%\setlist[enumerate, 2]{label=\asbuk*),ref=\asbuk*}

%% эконометрические сокращения
\DeclareMathOperator{\Cov}{Cov}
\DeclareMathOperator{\Corr}{Corr}
\DeclareMathOperator{\Var}{Var}
\DeclareMathOperator{\E}{E}
\def \hb{\hat{\beta}}
\def \hs{\hat{\sigma}}
\def \htheta{\hat{\theta}}
\def \s{\sigma}
\def \hy{\hat{y}}
\def \hY{\hat{Y}}
\def \v1{\vec{1}}
\def \e{\varepsilon}
\def \he{\hat{\e}}
\def \z{z}
\def \hVar{\widehat{\Var}}
\def \hCorr{\widehat{\Corr}}
\def \hCov{\widehat{\Cov}}
\def \cN{\mathcal{N}}


\begin{document}

\begin{enumerate}

	\item James Bond estimates the model $y_i = \beta x_i + u_i$ with OLS.
	We know that $u_i$ are normal, $\cN(0; \sigma^2)$ and independent.
	We also know that $x_1 = 2$ and $x_i = 1$ for $i\geq 2$.
	    \begin{enumerate}
      \item Find expected value and variance of $\hat \beta$.
      \item Is $\hat \beta$ efficient amond all linear unbiased estimators? Why or why not?
			\item Calculate probability limit of $\hat \beta$. Is $\hat \beta$ consistent?
    \end{enumerate}

  \item James Bond uses robust to heteroskedasticity White standard errors
	in the case of homoskedastic model $y_i = \beta x_i + u_i$, where
	$u_i \sim \cN(0; \sigma^2)$ and $u_i$ are independent.
	We also know that $x_1 = 2$ and $x_i = 1$ for $i\geq 2$.
 \begin{enumerate}
	 \item Calculate $se_{HC}(\hat\beta)$;
 	  \item Will $\hat \beta$ remain unbiased, efficient among all linear unbiased estimators? Why or why not?
		\item Is it correct to test significance of coefficients using $t$-distribution? How many degrees of freedom for $t$ distribution should James use? If it is not correct then what should James do?
 \end{enumerate}

 \item James Bond assumes that observations $y_i$ are independent and probability distribution for $y_i$ is given by the table

	 \begin{tabular}{cccc}
		 \toprule
		 $y_i$ & $y_i = 1$ & $y_i = 2$ & $y_i = 3$ \\
		 \midrule
		 probability & $3a$ & $2a$ & $1 - 5a$ \\
		 \bottomrule
	 \end{tabular}

	 James Bond has collected 400 observations: $100$ observations are equal to $1$, $100$ observations equal to $2$, and $200$ observations equal to $3$.

\begin{enumerate}
 \item Estimate unknown $a$ using maximum likelihood;
 \item Construct 95\% confidence interval for $a$.
\end{enumerate}


\item Using 400 observation James Bond estimated the logistic regression for probability
of passing econometrics exam. Predictors are: $x_i$ — number of study hours and $d_i$ —
dummy variable for having at least one dinner in nearby Khatchapuri restaurant:
\[
\hat \P(\text{pass}) = \Lambda(0.15 + 0.02x_i - 0.1 d_i),
\]

The estimate of coefficient covariance matrix is:
\[
\begin{pmatrix}
0.0004 & -0.0001 & 0 \\
-0.0001 & 0.0001 & 0 \\
0 & 0 & 0.0009 \\
\end{pmatrix}
\]

\begin{enumerate}
  \item Test whether the number of study hours influences the probability of passing the exam.
  \item Construct 95\% confidence interval for the difference of probabilities of passing the exam
	for two students. We know that both students have studied 20 hours, but only one of the students
	had a dinner in Khatchapuri restaurant.
  \item For which value of $d_i$ the marginal effect of increasing $x_i$ is bigger for $x_i = 30$?
\end{enumerate}

\item James Bond would like to check whether the econometrics final exam score, $y_i$, varies for
students that had at least one dinner at Khatchapuri and those who had not. Once again
$x_i$ is the number of study hours.

He estimates the same model three datasets:


\begin{tabular}{cccc}
	\toprule
	dataset & equation & $RSS$ & observations \\
	all students & $\hat y_i = 40 + 3x_i$ & 8000 & 400 \\
	at least one Khatchapuri dinner & $\hat y_i = 50 + 2x_i$ & 3000 & 200 \\
	no Khatchapuri dinner & $\hat y_i = 37 + 4x_i$ & 2000 & 200 \\
	\bottomrule
\end{tabular}

\begin{enumerate}
	\item Assume $d_i$ is those who have visited Khatchapuri at least once.
	What estimates will John obtain in the regression
	\[
    \hat y_i = \hat \beta_1 + \hat \beta_2 x_i + \hat \beta_3 d_i + \hat\beta_4 d_i x_i
	\]
	\item Using appropriate test check the econometrics final exam score varies for two type of students.
\end{enumerate}


$F_{a,b}$ distribution 5\% critical values are given below:

\begin{tabular}{cccc}
	\toprule
	& $b=100$ & $b=200$ & $b=500$  \\
	\midrule
	$a=1$ & $3.94$ & $3.89$ & $3.86$  \\
	$a=2$ & $3.09$ & $3.04$ & $3.01$ \\
	$a=3$ & $2.70$ & $2.65$ & $2.62$ \\
	$a=200$ & $1.34$ & $1.26$ & $1.21$ \\
	\bottomrule
\end{tabular}


\item The last problem :)

James Bond would like to estimate the model $y_i = \beta_1 + \beta_2 x_i + u_i$,
where $y_i$ — is the econometrics final exam result, $x_i$ is the number of study hours.

The problem is that James has only data on self-reported $x^*_i = x_i \cdot f_i$ where
$f_i$ is the famous Russian exageration factor. We know that $x_i \sim \cN(10;9)$,
$u_i \sim \cN(0;1)$, $f_i \sim \cN(3;6)$. Variables $(x_i)$, $(u_i)$ and $(f_i)$
are independent.

James Bond uses the OLS regression $\hat y_i = \hat \beta_1 + \hat\beta_2 x_i^*$.

\begin{enumerate}
	\item Find probability limit of $\hat\beta_2$. Is $\hat \beta_2$ consistent?
	\item Briefly describe the possible way to obtain consistent estimator.
\end{enumerate}





\end{enumerate}

\end{document}
